Fast Random Integer Generation in an Interval

### 2024-02-11 Presentation/unbiased.tex
remove FDR + summary table (need to fix vertical line)

\subsection{The OpenBSD Algorithm}
\begin{frame}{The OpenBSD Algorithm}
    \pause 
    \begin{itemize}[<+->]
        \item We shift the \ER{rejection interval} to the left:
        \onslide<+->{\begin{align*}\everymath{} % Allows specific coloring
            \overbrace{\ER{\underbrace{0,1,\ldots,\RNW - 1}_{\text{$\RNW$ values}}},\underbrace{\overbrace{\RNW,\ldots,n - 1,0,\ldots,\RNW - 1}^{\text{$n$ values}},\ldots,\overbrace{\RNW,\ldots,n - 1,0,\ldots,\RNW - 1}^{\text{$n$ values}}}_{\text{$\left(2^W \div n\right) \cdot n$ values}}}^{\text{$2^W$ values}}
        \end{align*}}
        \item Generate a uniform random number $x \in [0,2^W)$ until $x \geq \RNW$
        \item Return $x \Mod n$
    \end{itemize}

@@

%%% Array Drawing with Tikz instead of Equation
\iffalse
\onslide<+->{\begin{center}
    \begin{tikzpicture}[scale = 0.7]
        \draw[draw = goetheblau] (0.0,0.0) rectangle ++(5.0,1.0);
        \draw[draw = goetheblau,overbrace style] (0.0,1.0) -- (5.0,1.0) node [overbrace text style] {\GB{$n$ values}};
        \draw[draw = goetheblau] (5.0,0.0) rectangle ++(5.0,1.0);
        \draw[draw = goetheblau,overbrace style] (5.0,1.0) -- (10.0,1.0) node [overbrace text style] {\GB{$n$ values}};
        \node at (10.5,0.5) {$\cdots$};
        \draw[draw = goetheblau] (11.0,0.0) rectangle ++(5.0,1.0);
        \draw[draw = goetheblau,overbrace style] (11.0,1.0) -- (16.0,1.0) node [overbrace text style] {\GB{$n$ values}};
        \draw[draw = emorot] (16.0,0.0) rectangle ++(3.0,1.0);
        \draw[draw = emorot,overbrace style] (16.0,1.0) -- (19.0,1.0) node [overbrace text style] {$\text{\ER{\RNW values}}$};
    \end{tikzpicture}
\end{center}}
\fi

\subsection{The Java Algorithm}
\begin{frame}{The Java Algorithm}
    \pause 
    \begin{itemize}
        \item<1-> Consider $x - \left(x \Mod n\right)$ for $x \in [0,2^W)$: \only<3>{\begin{align*}\everymath{} % Allows specific coloring
            \overbrace{\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}},\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}},\ldots,\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}}}^{\text{$\left(2^W \div n\right) \cdot n$ values}},\ER{\overbrace{0,1,\ldots,\RNW - 1}^{\text{$\RNW$ values}}}
        \end{align*}} \only<4>{\begin{align*}\everymath{} % Allows specific coloring
            \overbrace{\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $0$}},\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}},\ldots,\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}}}^{\text{$\left(2^W \div n\right) \cdot n$ values}},\ER{\overbrace{0,1,\ldots,\RNW - 1}^{\text{$\RNW$ values}}}
        \end{align*}} \only<5>{\begin{align*}\everymath{} % Allows specific coloring
            \overbrace{\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $0$}},\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $n$}},\ldots,\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}}}^{\text{$\left(2^W \div n\right) \cdot n$ values}},\ER{\overbrace{0,1,\ldots,\RNW - 1}^{\text{$\RNW$ values}}}
        \end{align*}} \only<6>{\begin{align*}\everymath{} % Allows specific coloring
            \overbrace{\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $0$}},\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $n$}},\ldots,\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\substack{\text{mapped to} \\ 2^W - n - \RNW}}}^{\text{$\left(2^W \div n\right) \cdot n$ values}},\ER{\overbrace{0,1,\ldots,\RNW - 1}^{\text{$\RNW$ values}}}
        \end{align*}} \only<7->{\begin{align*}\everymath{} % Allows specific coloring
            \overbrace{\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $0$}},\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $n$}},\ldots,\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\substack{\text{mapped to} \\ 2^W - n - \RNW}}}^{\text{$\left(2^W \div n\right) \cdot n$ values}},\ER{\underbrace{\overbrace{\BB{0},1,\ldots,\RNW - 1}^{\text{$\RNW$ values}}}_{\substack{\text{mapped to} \\ 2^W - \RNW}}}
        \end{align*}}
        \item<8-> Only numbers in the \ER{leftover} interval get mapped to $\ER{2^W - \RNW} > 2^W - n$
        \item<9-> Algorithm: \begin{itemize}
            \item[(1)]<10-> Draw $x \in [0,2^W)$ and compute $r = x \Mod n$
            \item[(2)]<11-> Return $r$ if $x - r > 2^W - n$ else goto \textbf{\GB{(1)}}
        \end{itemize}
    \end{itemize}
\end{frame}

@@

\subsection{The Fast-Dice-Roller Algorithm}
\begin{frame}{The Fast-Dice-Roller Algorithm}
    \pause

@@

\subsection{The Bitmask Algorithm}
\begin{frame}{The Bitmask Algorithm}
    \pause

>>>>

\subsection{The OpenBSD Algorithm}\label{sec:2.1}
\begin{frame}{The OpenBSD Algorithm}
    \pause 
    \begin{itemize}[<+->]
        \item We shift the \ER{rejection interval} to the left:
        \onslide<+->{\begin{align*}\everymath{} % Allows specific coloring
            \overbrace{\ER{\underbrace{0,1,\ldots,\RNW - 1}_{\text{$\RNW$ values}}},\underbrace{\overbrace{\RNW,\ldots,n - 1,0,\ldots,\RNW - 1}^{\text{$n$ values}},\ldots,\overbrace{\RNW,\ldots,n - 1,0,\ldots,\RNW - 1}^{\text{$n$ values}}}_{\text{$\left(2^W \div n\right) \cdot n$ values}}}^{\text{$2^W$ values}}
        \end{align*}}
        \item Generate a uniform random number $x \in [0,2^W)$ until $x \geq \RNW$
        \item Return $x \Mod n$
    \end{itemize}

@@

\subsection{The Java Algorithm}\label{sec:2.2}
\begin{frame}{The Java Algorithm}
    \pause 
    \begin{itemize}
        \item<1-> Consider $x - \left(x \Mod n\right)$ for $x \in [0,2^W)$: \only<3>{\begin{align*}\everymath{} % Allows specific coloring
            \overbrace{\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}},\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}},\ldots,\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}}}^{\text{$\left(2^W \div n\right) \cdot n$ values}},\ER{\overbrace{0,1,\ldots,\RNW - 1}^{\text{$\RNW$ values}}}
        \end{align*}} \only<4>{\begin{align*}\everymath{} % Allows specific coloring
            \overbrace{\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $0$}},\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}},\ldots,\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}}}^{\text{$\left(2^W \div n\right) \cdot n$ values}},\ER{\overbrace{0,1,\ldots,\RNW - 1}^{\text{$\RNW$ values}}}
        \end{align*}} \only<5>{\begin{align*}\everymath{} % Allows specific coloring
            \overbrace{\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $0$}},\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $n$}},\ldots,\overbrace{0,1,\ldots,n - 1}^{\text{$n$ values}}}^{\text{$\left(2^W \div n\right) \cdot n$ values}},\ER{\overbrace{0,1,\ldots,\RNW - 1}^{\text{$\RNW$ values}}}
        \end{align*}} \only<6>{\begin{align*}\everymath{} % Allows specific coloring
            \overbrace{\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $0$}},\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $n$}},\ldots,\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\substack{\text{mapped to} \\ 2^W - n - \RNW}}}^{\text{$\left(2^W \div n\right) \cdot n$ values}},\ER{\overbrace{0,1,\ldots,\RNW - 1}^{\text{$\RNW$ values}}}
        \end{align*}} \only<7->{\begin{align*}\everymath{} % Allows specific coloring
            \overbrace{\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $0$}},\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\text{mapped to $n$}},\ldots,\underbrace{\overbrace{\BB{0},1,\ldots,n - 1}^{\text{$n$ values}}}_{\substack{\text{mapped to} \\ 2^W - n - \RNW}}}^{\text{$\left(2^W \div n\right) \cdot n$ values}},\ER{\underbrace{\overbrace{\BB{0},1,\ldots,\RNW - 1}^{\text{$\RNW$ values}}}_{\substack{\text{mapped to} \\ 2^W - \RNW}}}
        \end{align*}}
        \item<8-> We map every number to the next-smallest multiple of $n$
        \item<9-> Only numbers in the \ER{leftover} interval get mapped to $\ER{2^W - \RNW} > 2^W - n$
        \item<10-> Algorithm: \begin{itemize}
            \item[(1)]<11-> Draw $x \in [0,2^W)$ and compute $r = x \Mod n$
            \item[(2)]<12-> Return $r$ if $x - r > 2^W - n$ else goto \textbf{\GB{(1)}}
        \end{itemize}
    \end{itemize}
\end{frame}

@@

\subsection{The Bitmask Algorithm}\label{sec:2.3}
\begin{frame}{The Bitmask Algorithm}
    \pause

@@

%%% Array Drawing with Tikz instead of Equation + FastDiceRoller
\iffalse
\onslide<+->{\begin{center}
    \begin{tikzpicture}[scale = 0.7]
        \draw[draw = goetheblau] (0.0,0.0) rectangle ++(5.0,1.0);
        \draw[draw = goetheblau,overbrace style] (0.0,1.0) -- (5.0,1.0) node [overbrace text style] {\GB{$n$ values}};
        \draw[draw = goetheblau] (5.0,0.0) rectangle ++(5.0,1.0);
        \draw[draw = goetheblau,overbrace style] (5.0,1.0) -- (10.0,1.0) node [overbrace text style] {\GB{$n$ values}};
        \node at (10.5,0.5) {$\cdots$};
        \draw[draw = goetheblau] (11.0,0.0) rectangle ++(5.0,1.0);
        \draw[draw = goetheblau,overbrace style] (11.0,1.0) -- (16.0,1.0) node [overbrace text style] {\GB{$n$ values}};
        \draw[draw = emorot] (16.0,0.0) rectangle ++(3.0,1.0);
        \draw[draw = emorot,overbrace style] (16.0,1.0) -- (19.0,1.0) node [overbrace text style] {$\text{\ER{\RNW values}}$};
    \end{tikzpicture}
\end{center}}

\subsection{The Fast-Dice-Roller Algorithm}
\begin{frame}{The Fast-Dice-Roller Algorithm}
    \pause 
    \begin{itemize}[<+->]
        \item Build up $x$ bit-by-bit using uniform random bits \GB{\texttt{flip()}}
        \item Keep track of upper bound $\mathcal{B}$ for number \onslide<+->{  $\longrightarrow \quad x \in [0,\mathcal{B})$}
        \item Repeat until $\mathcal{B} \geq n$ \begin{itemize}[<+->]
            \item if $x < n$, return $x$
            \item else decrease $x$ and $\mathcal{B}$ by $n$ (\ER{rejection})
        \end{itemize}
    \end{itemize}

\end{frame}
\fi
